C/O in Protoplanetary Disks: The Effect of Radial Drift and Viscous Accretion

### 2015-07-15 tex/draft/ms.tex
update

%\usepackage{pdfsync}
\usepackage{subeqnarray}
\usepackage{natbib}
\usepackage{color}

@@

%\shorttitle{Critical Core Mass at Wide Separations}
%\shortauthors{Piso et al.}

@@

\emgr{Paragraph 1: introduce the topic, context and its relevance in the big picture . Gas giants are important. The chemical composition of their atmospheres constraints their formation and evolution. Is is therefore important to understand the disk well enough to (1) predict what kind of planet compositions result from planet formation in different parts of the disk, and (2) backtrack the planet formation location based on planet composition.}

%Main sequence stars commonly host giant planets (refs). The chemical composition of gas giant atmospheres can provide important constraints on their formation, accretion and migration history.

\emgr{Paragraph 2: disks are complex, a lot of dynamical and chemical processes going on. However, We have detections of organic molecules. We have detections of snowlines. We see complex chemistry and many molecules we can study.}

In recent years, the onset and development of sensitive infrared and (sub)millimeter spectroscopic observations has facilitated the detection of organic molecules in the outer regions of protoplanetary disks (e.g., \citealt{oberg10}, \citealt{oberg11b}, \citealt{oberg11c}, ...). Of particular importance are volatile compounds, since the location of their snowlines determines their relative abundance in gaseous and solid form in the protoplanetary disk, and thus the chemical composition of nascent giant planets. \emgr{Add part about snowlines, complex chemistry, many molecules.}

@@

\emgr{Start with the fact that we can't see C/O in disks directly. Then transition to planet atmospheres and modify the next sentence to follow up logically.} Notably, an important signature of giant planets atmospheric chemistry is the carbon to oxygen (C/O) ratio. Spectroscopic observations of gas giants such as WASP-12b have found atmospheric C/O ratios close to unity, substantially different from the Solar value of 0.54 \citep{madhu11}. One explanation for this discrepancy was proposed by \citet{oberg11}, who considered the fact that the main carries of carbon and oxygen, i.e. H$_2$O, CO$_2$ and CO, have different condensation temperatures. This changes the relative abundance of C and O in gaseous and solid form as a function of the snowline location of the volatiles mentioned above. \citet{oberg11} calculated analytically the C/O ratio in gas in dust as a function of semimajor axis for passive protoplanetary disks and reproduced a gas C/O ratio of order unity between the CO$_2$ and CO snowlines, where oxygen gas is highly depleted.

\emgr{Paragraph 4: the static disk model is a great start, but there are additional dynamical and chemical processes to take into account. Several studies have done that in some form or another --- cite Ali-Dib+14, Madhusudhan+14, Thiabaud+15. Briefly discuss their assumptions.}

@@

\textbf{Actively accreting disk with passive temperature profile.} We model the active disk as a thin disk with an $\alpha$-viscosity prescription \citep{shakura73}:
\begin{equation}
\label{eq:nu}
\nu=\alpha c_{\rm d} H_{\rm d}.
\end{equation}
Here $\nu$ is the kinematic viscosity, $\alpha < 1$ is a dimensionless coefficient and we choose $\alpha=0.01$, and $c_{\rm d}$, $H_{\rm d}$ are the isothermal sound speed and disk scale height, respectively:
\begin{subeqnarray}
\label{eq:cdHd}
c_{\rm d} &=& \sqrt{\frac{k_{\rm B} T_{\rm d}}{\mu m_{\rm p}}} \slabel{eq:cd} \\
H_{\rm d}&=& \frac{c_{\rm d}}{\Omega_{\rm k}} \slabel{eq:Hd},
\end{subeqnarray}
where $k_{\rm B}$ is the Boltzmann constant, $\mu$ is the mean molecular weight of the gas, $m_{\rm p}$ is the proton mass, and $\Omega_{\rm k}$ is the Keplerian angular velocity, $\Omega_{\rm k} \equiv \sqrt{G M_*/r^3}$, with $G$ the gravitational constant and $M_*$ the stellar mass. We choose $M_*=M_{\odot}$ and $\mu=2.35$, corresponding to the Solar composition go hydrogen and helium. The temperature profile for an active disk is assumed to be the same as for the passive disk and given by Equation (\ref{eq:diskT}). From Equations (\ref{eq:nu}) and (\ref{eq:cdHd}), the viscosity can thus be expressed as a power-law in radius, $\nu \propto r^{\gamma}$, with $\gamma=15/14 \approx 1$ for our choice of parameters. Following \citet{armitage10}, we define $R \equiv r/r_{\rm c}$ and $\nu_{\rm c} \equiv \nu(r_{\rm c})$, where $r_{\rm c}$ is a characteristic disk radius. We choose $r_{\rm c}=100$ AU. The gas surface density is then given by the self-similar solution
\begin{equation}
\label{eq:Sigmaact}
\Sigma_{\rm d}(R, T) = \frac{M_{\rm d}}{2 \pi r_{\rm c}^2 R^{\gamma}} T^{(-5/2-\gamma)/(2-\gamma)} \exp{\Big[-\frac{R^{-(2-\gamma)}}{T}}\Big],
\end{equation}
where $M_{\rm d}$ is the total disk mass and
\begin{subeqnarray}
\label{eq:T}
T & \equiv & \frac{t}{t_{\rm c}} + 1 \\
t_{\rm c} & \equiv & \frac{1}{3(2-\gamma)} \frac{r_{\rm c}^2}{\nu_{\rm c}},
\end{subeqnarray}
where $t$ is time. We take $M_{\rm d}=0.1 M_{\odot}$ (e.g., \citealt{birnstiel12}), but note that our results are insensitive to this choice (see Section \ref{sec:discussion}).

\textbf{Active disk steady-state solution.} Calculating the midplane temperature self-consistently for an active disk is non-trivial, so instead we use the steady state solution for the surface density and temperature for a thin disk (e.g., \citealt{armitage10}):

\begin{subeqnarray}
\label{eq:activeT}
T_{\rm d}^4 & = & \frac{3 G M_* \dot{M}}{8 \pi \sigma r^3} \slabel{eq:Tdact} \\
\dot{M} & = & 3 \pi \nu \Sigma_{\rm d} \slabel{eq:Mdot},
\end{subeqnarray}
where $\dot{M}$, the mass flux, is constant for steady-state \footnote{Equations (\ref{eq:Tdact}) and (\ref{eq:Mdot}) are valid when $r \gg R_*$, the stellar radius, which is the regime in which we conduct our study.}. Based on disk observations (e.g., \citealt{andrews10}), we choose $\dot{M}=10^{-8} M_{\odot}$ yr$^{-1}$. We can then easily determine the radial temperature profile from Equation (\ref{eq:Tdact}), $c_{\rm d}$ and $H_{\rm d}$ from Equation (\ref{eq:cdHd}), as well as the viscosity $\nu$ from Equation (\ref{eq:nu}) for a given $\alpha$. For consistency, we choose  $\alpha=0.01$ as in the previous case. Finally, we determine $\Sigma_{\rm d}$ from Equation (\ref{eq:Mdot}).

\textbf{Static disk.} For comparison purposes with \citet{oberg11}, we also use a static disk model, with $\Sigma_{\rm d}$ and $T_{\rm d}$ described by Equations (\ref{eq:disksigma}) and (\ref{eq:diskT}). The static model does not take into account gas accretion on to the central star or radial drift of solids (see Section \ref{sec:drift}).

@@

For simplicity purposes, we calculate the radial drift timescale, $t_{\rm r, drift}$, for a passive disk in this section, but most of our conclusions hold true for an active disk as well. Figure \ref{fig:timescales} shows $t_{\rm des}$, $t_{\rm r, drift}$ and $t_{\rm gas, acc}$ as a function of particle size at three different locations in the disk, corresponding to the H$_2$O, CO$_2$ and CO snowlines in the static disk. As expected, micron-sized particles desorb on very short timescales of $\sim 1-1000$ years in the close vicinity of their respective snowlines, since the desorption rate depends exponentially on temperature and hence disk location (see Equation \ref{eq:Rdes}).  On the other hand, their radial drift timescale exceeds the typical disk lifetime of a few Myr by several orders of magnitude due to their strong coupling with the gas. Thus for small particles in a passive disk, the snowline locations and the C/O ratio are the same as for a static disk (see Figure 1 from \citealt{oberg11}) \footnote{This is not true for an active disk, however, where gas accretion causes even micron-sized particles to drift significantly before desorbing, as we show in section \ref{sec:snowlines}.}. At the other extreme, kilometer-sized particles are unaffected by gas drag and have long desorption timescales ($\gg$1 Myr ), and the snowline locations and C/O ratio remain unchanged in this case as well. This is true for both passive and active disks, since large planetesimals are decoupled from the gas and hence unaffected by gas accretion onto the host star.

@@

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{../../figs/drift_timescales_betaS1_gas_acc_new.pdf}
%\vspace{-0.5in}
\caption{Relevant timescales for dynamical effects in the desorption process: $t_{\rm r, drift}$ (solid lines), $t_{\rm evap}$ (dashed lines) and $t_{\rm gas, acc}$ (dotted lines). The timescales are calculated at three representative locations in the disk, i.e. the H$_2$O, CO$_2$ and CO snowlines. For our choice of parameters, the snowlines are located at $\sim$0.7 AU (blue lines), $\sim$8.6 AU (green lines) and $\sim$59 AU (red lines), respectively. The horizontal dashed line represents a typical disk lifetime of 3 Myr. Radial drift and gas accretion affect desorption in the regions where their respective timescales, i.e. $t_{\rm r, drift}$ and $t_{\rm gas, acc}$, are comparable to the desorption timescale $t_{\rm des}$.} 
\label{fig:timescales}
\end{figure}

@@

\begin{figure*}[t!]
\centering
\includegraphics[width=0.7\textwidth]{../../figs/desorption_distance_actual_vs_estimated_passive_active.pdf}
%\vspace{-0.5in}
\caption{Desorption distance estimated from analytic calculations (see text) as a function of the desorption distance calculated numerically, for the range of particle sizes that desorb at a fixed distance regardless of initial location (see Figure \ref{fig:snowlines} and text). The estimate is performed for a passive disk (left panels) and an active disk (right panels).  The particles are composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels). The analytic approximation is in good agreement with the numerical result for most cases. \textbf{Figure belongs to section 3.}}  %  (See text for a description of evolution to yet higher masses.) 
\label{fig:an_vs_actual}
\end{figure*}

@@

As we show in Section \ref{sec:COratio}, a drifting particle that desorbs will do so almost instantaneously and will lose most of its mass very close to the distance at which it fully desorbs. Thus a particle's final location will depend on whether a grain initially at a certain distance is completely desorbed or not within the disk lifetime of 3 Myr. For example, larger grains take longer to desorb and hence are more likely to not evaporate fully in a given timeframe.

Figure \ref{fig:snowlines} shows our results for H$_2$O, CO$_2$ and CO particles, and for both a passive and an active disk. We do not show the equivalent result for the steady-state active disk since the trends are qualitatively the same as for the active disk with the temperature profile given by Equation (\ref{eq:diskT}). Kilometer-sized bodies do not drift or desorb during the disk lifetime both for a passive and an active disk. Similarly, micron- to mm-sized particles neither drift nor desorb unless they are located inside the static disk snowlines. This is not in contradiction with the desorption timescales for small particles from Figure \ref{fig:timescales}, since small grains only desorb fast at or nearby their respective snowlines, as mentioned in Section \ref{sec:timescales}.  In an active disk, however, micron-to mm-sized grains do drift significantly since they move at the same velocity as the accreting gas. For $0.5$ cm $\lesssim s_0 \lesssim$ 700 cm in a passive disk and $0.001$ cm $\lesssim s_0 \lesssim$ 700 cm in an active disk, we notice that particles of initial size $s_0$ desorb at a fixed distance $r_{\rm des}$ regardless of their original location in the disk. In fact, the only grains that will both drift and evaporate are those that reach their fixed final location (represented by the horizontal curves in Figure \ref{fig:snowlines}) within the disk lifetime. We show in section \ref{sec:COratio} that this result is essential in determining the C/O ratio throughout the disk for different particle sizes.

\begin{figure*}[t!]
\centering
\includegraphics[width=0.8\textwidth]{../../figs/s_t_a.pdf}
%\vspace{-0.5in}
\caption{Left panels: size of desorbing H$_2$O particles as a function of time, for different initial particle sizes and for three initial locations in a passive disk: 1 AU (top left), 10 AU (middle left) and 100 AU (bottom left). Particles desorb almost instantaneously. Right panel: fractional mass lost by the desorbing particles as a function of the particle's location as it drifts, for different initial particle sizes, and at the same initial locations presented in the left panel. Particles lose most of their mass very close to the distance at which they fully desorb.}
\label{fig:s_t_a}
\end{figure*}

@@

We now use our model and the results of Section \ref{sec:snowlines} to determine the H$_2$O, CO$_2$ and CO snowline locations and the C/O ratio in disks with static chemistry that experience radial drift of solids and gas accretion on to the central star. In Section \ref{sec:snowlines} we showed that solid particles which drift and fully desorb during the lifetime of the protoplanetary disk will do so (1) instantaneously and at (2) at a fixed stellocentric distance, regardless of their initial location in the disk. Figure \ref{fig:s_t_a} confirms both of these effects. The left panels show the size evolution with time for H$_2$O particles of various initial sizes, starting at three different initial locations in a passive disk \footnote{Our conclusions remain valid for an active disk and for particles composed of CO$_2$ or CO.}.  Indeed, the solid H$_2$O particles evaporate almost instantly, although the time $t_{\rm des}$ at which a particle of a given initial size desorbs depends on its initial distance. A particle located at the initial time $t_0$ at a distance such that $t_{\rm des}>t_{\rm d}$ will therefore not desorb during the disk lifetime. However, our model assumes that particles drift continuously at any location in the disk. Therefore, a particle that can fully desorb during the disk lifetime for at least one initial location will always desorb, at a fixed distance as discussed in Section \ref{sec:snowlines} and displayed in Figure \ref{fig:snowlines}. The right panels of Figure \ref{fig:s_t_a} show that the drifting grains lose most of their mass in a very narrow distance range; moreover, this distance is the same for a given initial particle size, no matter where the particle started drifting at the time $t_0$ when the simulation is started. Figure \ref{fig:s_t_a} thus proves claims (1) and (2) above. It follows that the H$_2$O, CO$_2$ and CO snowlines are fixed for a given initial particle size and disk model (passive or active). The C/O ratio will then only depend on disk properties, grain size, and the abundance of H$_2$O, CO$_2$ and CO relative to the H$_2$ abundance in the disk midplane.

@@

We use the relative number densities of C and O in their different molecular forms (H$_2$O, CO$_2$ and CO) from Table 1 of \citet{oberg11}. \emgr{Is it necessary to reproduce that table? Seems sort of redundant.}. Figure \ref{fig:CO_ratio} shows the C/O ratio in gas and dust as a function of semimajor axis for a passive disk, an active disk with a passive temperature profile, and a steady-state active disk \emgr(not yet there). The C/O ratio for a static disk is shown as a guidline. The plot is consistent with Figure \ref{fig:snowlines}. For the passive disk, only grains larger than $\sim$0.5 cm drift, desorb and thus form a snowline. In contrast, even $\sim$micron-sized grains drift and desorb for the active disk, since the flow towards the host star together with the accreting gas. For the same particle size, the snowline locations are slightly closer to the central star in the active disk, due to the fact that the accreting gas adds an additional component to the drift velocity of the solids (cf. Equation \ref{eq:rdotact}). Perhaps the most interesting feature is the fact that the snowlines are pushed inwards as the grain size increases. While the plot only shows the snowlines and C/O ratio for particle sizes up to $\sim$7 m, we have found that even kilometer sized boulders are able to drift and desorb for both the passive and the active disk. However, bodies larger than $\sim$7 m will evaporate at the same location as the meter-sized planetesimals. Thus the innermost snowlines (depicted in red in Figure \ref{fig:CO_ratio}) set the limit on how far close in the H$_2$O, CO$_2$ and CO snowlines can be pushed due to radial drift and gas accretion on to the host star. Realistic grain size distributions in disks are dominated by large grains (e.g., \citealt{dalessio01}, \citealt{birnstiel12}). Therefore, the snowlines produced by the largest drifting solids in our model set the inner limit on the snowline locations and are insensitive to a particular grain size distribution. \emgr{There will be 1-2 additional sentences for the third panel, the steady-state disk, once that finishes running.}

@@

\begin{figure*}[t!]
\centering
\includegraphics[width=\textwidth]{../../figs/time_plots.pdf}
%\vspace{-0.5in}
\caption{Desorption / final distance as a function of initial position in the disk for particles of initial size $s_0=1$ m (left panels) and $s_0=10$ m (right panels), for grains composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels). The evolution is shown at four representative timescales: $10^4$ yr (cyan curve), $10^5$ yr (red curve), 1 Myr (green curve), and 3 Myr, the disk lifetime (blue curve). For a given particle size, the desorption distance, and hence the H$_2$O, CO$_2$ and CO snowlines, have the same location regardless of the timescale at which the simulation is stopped.}
\label{fig:timeplots}
\end{figure*}

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{../../figs/desorption_distance_varying_Md.pdf}
%\vspace{-0.5in}
\caption{Desorption / final distance as a function of initial position in the disk for H$_2$O particles of initial size of 1 m, for total disk masses $M_{\rm d}=0.1 M_{\odot}$ (solid lines) and $M_{\rm d}=0.01 M_{\odot}$ (dashed lines). The timescales of the simulations are the same as in Figure \ref{fig:timeplots}. A lower disk mass does not change the snowline location.}
\label{fig:varMd}
\end{figure}

@@

Our goals in this paper were (1) to gain a detailed qualitative and quantitative understanding of the effect or radial drift and gas accretion on to the central star on snowline locations and the C/O ratio in disks, and (2) to obtain a limit on how close in the snowlines can be pushed due to drift and gas accretion. We have thus used a simplified model and out of necessity neglected potentially significant dynamical and chemical effects. In what follows, we discuss these limitations and their effects. We note that our future work will address some of these issues.

>>>>

%\usepackage{pdfsync}
\usepackage{subeqnarray}
\usepackage{natbib}
\usepackage{color}
\usepackage[utf8]{inputenc}

@@

@@

\emgr{Paragraph 1: introduce the topic, context and its relevance in the big picture. Gas giants are important. The chemical composition of their atmospheres constraints their formation and evolution. Is is therefore important to understand the disk well enough to (1) predict what kind of planet compositions result from planet formation in different parts of the disk, and (2) backtrack the planet formation location based on planet composition.}

Main sequence stars commonly host giant planets (refs). The chemical composition of gas giant atmospheres can provide important constraints on their formation, accretion and migration history. \emgr{Add the rest as outlined above.} \\

\emgr{Paragraph 2: disks are complex, a lot of dynamical and chemical processes going on. However, we have detections of organic molecules. We have detections of snowlines. We see complex chemistry and many molecules we can study.}

In recent years, the onset and development of sensitive infrared and (sub)millimeter spectroscopic observations has facilitated the detection of organic molecules in the outer regions of protoplanetary disks (e.g., \citealt{oberg10}, \citealt{oberg11b}, \citealt{oberg11c}, ...). Of particular importance are volatile compounds, since the location of their snowlines determines their relative abundance in gaseous and solid form in the protoplanetary disk, and thus the chemical composition of nascent giant planets. \emgr{Add part about snowlines, complex chemistry, many molecules.} \\

@@

\emgr{Start with the fact that we can't see C/O in disks directly. Then transition to planet atmospheres and modify the next sentence to follow up logically.} Notably, an important signature of giant planets atmospheric chemistry is the carbon to oxygen (C/O) ratio. Spectroscopic observations of gas giants such as WASP-12b have found atmospheric C/O ratios close to unity, substantially different from the Solar value of 0.54 \citep{madhu11}. One explanation for this discrepancy was proposed by \citet{oberg11}, who considered the fact that the main carries of carbon and oxygen, i.e. H$_2$O, CO$_2$ and CO, have different condensation temperatures. This changes the relative abundance of C and O in gaseous and solid form as a function of the snowline location of the volatiles mentioned above. \citet{oberg11} calculated analytically the C/O ratio in gas in dust as a function of semimajor axis for passive protoplanetary disks and reproduced a gas C/O ratio of order unity between the CO$_2$ and CO snowlines, where oxygen gas is highly depleted. \\

\emgr{Paragraph 4: the static disk model is a great start, but there are additional dynamical and chemical processes to take into account. Several studies have done that in some form or another --- cite Ali-Dib+14, Madhusudhan+14, Thiabaud+15. Briefly discuss their assumptions.} \\

@@

\textbf{Actively accreting disk with passive temperature profile.} We model the active disk as a thin disk with an $\alpha$-viscosity prescription \citep{shakura73}:
\begin{equation}
\label{eq:nu}
\nu=\alpha c_{\rm d} H_{\rm d}.
\end{equation}
Here $\nu$ is the kinematic viscosity, $\alpha < 1$ is a dimensionless coefficient and we choose $\alpha=0.01$, and $c_{\rm d}$, $H_{\rm d}$ are the isothermal sound speed and disk scale height, respectively:
\begin{subeqnarray}
\label{eq:cdHd}
c_{\rm d} &=& \sqrt{\frac{k_{\rm B} T_{\rm d}}{\mu m_{\rm p}}} \slabel{eq:cd} \\
H_{\rm d}&=& \frac{c_{\rm d}}{\Omega_{\rm k}} \slabel{eq:Hd},
\end{subeqnarray}
where $k_{\rm B}$ is the Boltzmann constant, $\mu$ is the mean molecular weight of the gas, $m_{\rm p}$ is the proton mass, and $\Omega_{\rm k} \equiv \sqrt{G M_*/r^3}$ is the Keplerian angular velocity,  with $G$ the gravitational constant and $M_*$ the stellar mass. We choose $M_*=M_{\odot}$ and $\mu=2.35$, corresponding to the Solar composition of hydrogen and helium. The temperature profile for an active disk is assumed to be the same as for the passive disk and given by Equation (\ref{eq:diskT}). From Equations (\ref{eq:nu}) and (\ref{eq:cdHd}), the viscosity can thus be expressed as a power-law in radius, $\nu \propto r^{\gamma}$, with $\gamma=15/14 \approx 1$ for our choice of parameters. Following \citet{armitage10}, we define $R \equiv r/r_{\rm c}$ and $\nu_{\rm c} \equiv \nu(r_{\rm c})$, where $r_{\rm c}$ is a characteristic disk radius. We choose $r_{\rm c}=100$ AU. The gas surface density is then given by the self-similar solution
\begin{equation}
\label{eq:Sigmaact}
\Sigma_{\rm d}(R, T) = \frac{M_{\rm d}}{2 \pi r_{\rm c}^2 R^{\gamma}} T^{(-5/2-\gamma)/(2-\gamma)} \exp{\Big[-\frac{R^{-(2-\gamma)}}{T}}\Big],
\end{equation}
where $M_{\rm d}$ is the total disk mass and
\begin{subeqnarray}
\label{eq:T}
T & \equiv & \frac{t}{t_{\rm c}} + 1 \\
t_{\rm c} & \equiv & \frac{1}{3(2-\gamma)} \frac{r_{\rm c}^2}{\nu_{\rm c}},
\end{subeqnarray}
where $t$ is time. We choose $M_{\rm d}=0.1 M_{\odot}$ (e.g., \citealt{birnstiel12}), but we note that our results are insensitive to this choice (see Section \ref{sec:discussion}).

\textbf{Active disk steady-state solution.} Calculating the midplane temperature self-consistently for an active disk is non-trivial, so instead we use the steady-state solution for the surface density and temperature for a thin disk (e.g., \citealt{armitage10}):

\begin{subeqnarray}
\label{eq:activeT}
T_{\rm d}^4 & = & \frac{3 G M_* \dot{M}}{8 \pi \sigma r^3} \slabel{eq:Tdact} \\
\dot{M} & = & 3 \pi \nu \Sigma_{\rm d} \slabel{eq:Mdot},
\end{subeqnarray}
where $\dot{M}$, the mass flux, is constant in steady-state \footnote{Equations (\ref{eq:Tdact}) and (\ref{eq:Mdot}) are valid when $r \gg R_*$, the stellar radius, which is the regime in which we conduct our study.}. Based on disk observations (e.g., \citealt{andrews10}), we choose $\dot{M}=10^{-8} M_{\odot}$ yr$^{-1}$. We can then easily determine the radial temperature profile from Equation (\ref{eq:Tdact}), $c_{\rm d}$ and $H_{\rm d}$ from Equation (\ref{eq:cdHd}), as well as the viscosity $\nu$ from Equation (\ref{eq:nu}) for a given $\alpha$. For consistency, we choose  $\alpha=0.01$ as in the previous case. Finally, we determine $\Sigma_{\rm d}$ from Equation (\ref{eq:Mdot}).

\textbf{Static disk.} To compare our results with those of \citet{oberg11}, we also use a static disk model, with $\Sigma_{\rm d}$ and $T_{\rm d}$ described by Equations (\ref{eq:disksigma}) and (\ref{eq:diskT}). The static model does not take into account gas accretion on to the central star or radial drift of solids (see Section \ref{sec:drift}).

@@

For simplicity purposes, we calculate the radial drift timescale, $t_{\rm r, drift}$, for a passive disk in this section, but most of our conclusions hold true for an active disk as well. Figure \ref{fig:timescales} shows $t_{\rm des}$, $t_{\rm r, drift}$ and $t_{\rm gas, acc}$ as a function of particle size at three different locations in the disk, corresponding to the H$_2$O, CO$_2$ and CO snowlines in the static disk. As expected, micron-sized particles desorb on very short timescales of $\sim 1-1000$ years in the close vicinity of their respective snowlines, since the desorption rate depends exponentially on temperature and hence on disk location (see Equation \ref{eq:Rdes}).  On the other hand, their radial drift timescale exceeds the typical disk lifetime of a few Myr by several orders of magnitude due to their strong coupling with the gas. Thus for small particles in a passive disk, the snowline locations and the C/O ratio are the same as for a static disk (see Figure 1 from \citealt{oberg11}) \footnote{This is not true for an active disk, however, where gas accretion causes even micron-sized particles to drift significantly before desorbing, as we show in section \ref{sec:snowlines}.}. At the other extreme, kilometer-sized particles are unaffected by gas drag and have long desorption timescales ($\gg$1 Myr ), and the snowline locations and C/O ratio remain unchanged in this case as well. This is true for both passive and active disks, since large planetesimals are decoupled from the gas and hence unaffected by gas accretion onto the host star.

@@

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{../../figs/drift_timescales_betaS1_gas_acc_new.pdf}
%\vspace{-0.5in}
\caption{Relevant timescales for dynamical effects in the desorption process: $t_{\rm r, drift}$ (solid lines), $t_{\rm des}$ (dashed lines) and $t_{\rm gas, acc}$ (dotted lines). The timescales are calculated at three representative locations, i.e. the H$_2$O, CO$_2$ and CO snowlines in the static disk. For our choice of parameters, the snowlines are located at $\sim$0.7 AU (blue lines), $\sim$8.6 AU (green lines) and $\sim$59 AU (red lines), respectively. The horizontal dashed line represents a typical disk lifetime of 3 Myr. Radial drift and gas accretion affect desorption in the regions where their respective timescales, i.e. $t_{\rm r, drift}$ and $t_{\rm gas, acc}$, are comparable to the desorption timescale $t_{\rm des}$.} 
\label{fig:timescales}
\end{figure}

@@

\begin{figure*}[t!]
\centering
\includegraphics[width=0.7\textwidth]{../../figs/desorption_distance_actual_vs_estimated_passive_active.pdf}
%\vspace{-0.5in}
\caption{Desorption distance estimated from analytic calculations (see text) as a function of the desorption distance calculated numerically, for the range of particle sizes that desorb at a fixed distance regardless of their initial location (see Figure \ref{fig:snowlines} and text). The estimate is performed for a passive disk (left panels) and an active disk (right panels).  The particles are composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels). The analytic approximation is in good agreement with the numerical result for most cases.}
\label{fig:an_vs_actual}
\end{figure*}

@@

As we show in Section \ref{sec:COratio}, a drifting particle that desorbs will do so almost instantaneously and will lose most of its mass very close to the distance at which it fully evaporates. Thus a particle's final location will depend on whether a grain initially at a certain distance is completely desorbed or not within the disk lifetime of 3 Myr. For example, larger grains take longer to desorb and hence are more likely to not evaporate fully in a given timeframe.

Figure \ref{fig:snowlines} shows our results for H$_2$O, CO$_2$ and CO particles, for both a passive and an active disk. We do not show the equivalent result for the steady-state active disk since the trends are qualitatively the same as for the active disk with the temperature profile given by Equation (\ref{eq:diskT}). Kilometer-sized bodies do not drift or desorb during the disk lifetime neither for a passive nor for an active disk. Similarly, micron- to mm-sized particles in the passive disk do not drift or desorb unless they are located inside the static snowlines. This is not in contradiction with the desorption timescales for small particles from Figure \ref{fig:timescales}, since small grains only desorb fast at or nearby their respective snowlines, as mentioned in Section \ref{sec:timescales}.  In an active disk, however, micron-to mm-sized grains do drift significantly since they move at the same velocity as the accreting gas. For $0.5$ cm $\lesssim s_0 \lesssim$ 700 cm in a passive disk and $0.001$ cm $\lesssim s_0 \lesssim$ 700 cm in an active disk, we notice that particles of initial size $s_0$ desorb at a fixed distance $r_{\rm des}$ regardless of their original location in the disk. In fact, the only grains that will both drift and evaporate are those that reach their fixed final location (represented by the horizontal curves in Figure \ref{fig:snowlines}) within the disk lifetime. We show in section \ref{sec:COratio} that this result is essential in determining the C/O ratio throughout the disk for different particle sizes.

\begin{figure*}[t!]
\centering
\includegraphics[width=0.8\textwidth]{../../figs/s_t_a.pdf}
%\vspace{-0.5in}
\caption{Left panels: size of desorbing H$_2$O particles as a function of time, for different initial particle sizes and for three initial locations in a passive disk: 1 AU (top left), 10 AU (middle left) and 100 AU (bottom left). Particles desorb almost instantaneously. Right panel: fractional mass of the desorbing particles as a function of the particle's location as it drifts, for different initial particle sizes, and at the same initial locations presented in the left panel. Particles lose most of their mass very close to the distance at which they fully desorb.}
\label{fig:s_t_a}
\end{figure*}

@@

We now use our model and the results of Section \ref{sec:snowlines} to determine the H$_2$O, CO$_2$ and CO snowline locations and the C/O ratio in disks with static chemistry that experience radial drift of solids and gas accretion on to the central star. In Section \ref{sec:snowlines}, we showed that solid particles that drift and fully desorb during the lifetime of the protoplanetary disk do so (1) instantaneously, and (2) at a fixed stellocentric distance, regardless of their initial location in the disk. Figure \ref{fig:s_t_a} confirms both of these effects. The left panels show the size evolution with time for H$_2$O particles of various initial sizes, starting at three different initial locations in a passive disk \footnote{Our conclusions remain valid for an active disk and for particles composed of CO$_2$ or CO.}.  Indeed, the solid H$_2$O particles evaporate almost instantly, although the time $t_{\rm des}$ at which a particle of a given initial size desorbs depends on its initial distance. A particle located at the initial time $t_0$ at a distance such that $t_{\rm des}>t_{\rm d}$ will therefore not desorb during the disk lifetime. However, our model assumes that particles drift continuously at any location in the disk. Therefore, a particle that can fully desorb during the disk lifetime for at least one initial location will always desorb, at a fixed distance as discussed in Section \ref{sec:snowlines} and displayed in Figure \ref{fig:snowlines}. The right panels of Figure \ref{fig:s_t_a} show that the drifting grains lose most of their mass in a very narrow distance range; moreover, this distance is the same for a given initial particle size, no matter where the particle started drifting at the time $t_0$ when the simulation is started. Figure \ref{fig:s_t_a} thus proves claims (1) and (2) above. It follows that the H$_2$O, CO$_2$ and CO snowlines are fixed for a given initial particle size and disk model (passive or active). The C/O ratio will then only depend on disk properties, grain size, and the abundance of H$_2$O, CO$_2$ and CO relative to the H$_2$ abundance in the disk midplane.

@@

We use the relative number densities of C and O in their different molecular forms (H$_2$O, CO$_2$ and CO) from Table 1 of \citet{oberg11}. \emgr{Is it necessary to reproduce that table? Seems a bit redundant.}. Figure \ref{fig:CO_ratio} shows the C/O ratio in gas and dust as a function of semimajor axis for a passive disk, an active disk with a passive temperature profile, and a steady-state active disk \emgr{(not yet there)}. The C/O ratio for a static disk is shown as a guideline. The plot is consistent with Figure \ref{fig:snowlines}. For the passive disk, only grains larger than $\sim$0.5 cm drift, desorb and thus form a snowline. In contrast, even $\sim$micron-sized grains drift and desorb for the active disk, since they flow towards the host star together with the accreting gas. For the same particle size, the snowline locations are slightly closer to the central star in the active disk, due to the fact that the accreting gas adds an additional component to the drift velocity of the solids (cf. Equation \ref{eq:rdotact}). Perhaps the most interesting feature is the fact that the snowlines are pushed inwards as the grain size increases. While the plot only shows the snowlines and C/O ratio for particle sizes up to $\sim$7 m, we have found that almost kilometer-sized boulders are able to drift and desorb for both the passive and the active disk. However, bodies larger than $\sim$7 m will evaporate at the same location as the meter-sized planetesimals. Thus the innermost snowlines (depicted in red in Figure \ref{fig:CO_ratio}) set the limit on how close in the H$_2$O, CO$_2$ and CO snowlines can be pushed due to radial drift and gas accretion on to the host star. Realistic grain size distributions in disks are dominated by large grains (e.g., \citealt{dalessio01}, \citealt{birnstiel12}). Therefore, the snowlines produced by the largest drifting solids in our model set the inner limit on the snowline locations and are insensitive to a particular grain size distribution. \emgr{There will be 1-2 additional sentences for the third panel, the steady-state disk, once that finishes running.}

@@

\begin{figure*}[t!]
\centering
\includegraphics[width=\textwidth]{../../figs/time_plots.pdf}
%\vspace{-0.5in}
\caption{Desorption / final distance as a function of initial position in the disk for particles of initial size $s_0=1$ m (left panels) and $s_0=10$ cm (right panels), for grains composed of H$_2$O (top panels), CO$_2$ (middle panels) and CO (bottom panels). The evolution is shown at four representative timescales: $10^4$ yr (cyan curve), $10^5$ yr (red curve), 1 Myr (green curve), and 3 Myr, the disk lifetime (blue curve). For a given particle size, the desorption distance, and hence the H$_2$O, CO$_2$ and CO snowlines, have the same location regardless of the time at which the simulation is stopped.}
\label{fig:timeplots}
\end{figure*}

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{../../figs/desorption_distance_varying_Md.pdf}
%\vspace{-0.5in}
\caption{Desorption / final distance as a function of initial position in the disk for H$_2$O particles of initial size of 1 m, for total disk masses $M_{\rm d}=0.1 M_{\odot}$ (solid lines) and $M_{\rm d}=0.01 M_{\odot}$ (dashed lines). The timescales of the simulations and their color code are the same as in Figure \ref{fig:timeplots}. A lower disk mass does not change the snowline location.}
\label{fig:varMd}
\end{figure}

@@

Our goals in this paper were (1) to gain a detailed qualitative and quantitative understanding of the effect of radial drift and gas accretion on to the central star on snowline locations and the C/O ratio in disks, and (2) to obtain a limit on how close in the snowlines can be pushed due to drift and gas accretion. We have thus used a simplified model and out of necessity neglected potentially significant dynamical and chemical processes. In what follows, we discuss these limitations and their effects. We note that our future work will address some of these issues.
